---
title: "Reproducible Research: Peer Assessment 2"
author: "Paul Engler"
date: "July 17, 2015"
output: 
   html_document:
      keep_md: true
---

## Summary/Synopsis
- Synopsis: Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences.

## Data Processing

- There should be a section titled Data Processing which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the cache = TRUE option for certain code chunks

- Set up the environment for the data analysis
- Retreive and uncompress the data file

```{r Data_Retreival, echo=TRUE}
dataURL<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
dataFile <-"repdata.data.StormData.csv.bz2"
baseDir<-getwd()
dataDir<-paste(baseDir,"data",sep="/")
dataFile<-paste(dataDir,dataFile,sep="/")
dataFileExt <- "repdata.data.StormData.csv"
dataFileExt  <- paste(dataDir,dataFileExt,sep="/")

# Only download the and unpack the file if it does not exist
# This should be a one time performance penalty on the first execution
# and will leave the uncompress data on disk so that we can be more
# efficient in reading the data later on. 
if (!file.exists(dataFile)) { 
   dir.create(dataDir,showWarnings = FALSE)
   download.file(url=dataURL,destfile=dataFile,method="curl")
}   

# Check if we have the uncompressed file
if (!file.exists(dataFileExt)) {
   # Assumes we are on a Linux system with bunzip2 in the path
   # uncompress the file and leave the compressed file in place
   system(paste("bunzip2 -k", dataFile))
}
```

- Read in the data file
```{r Read_Data, echo=TRUE}
# The data should all be in an uncompress format on disk
# we can now read in the data using read.csv and
# defining our colClasses to minimize our in-memory footprint 

# The data set has 37 columns,
importColumns <- rep("NULL",37)

# Populate the colClasses vectors will NULLs (will import nothing)
# we are interested in the following columns
#importColumns[2] =  "character"    #BGN_DATE
importColumns[8] =  "factor"       #EVTYPE
importColumns[23] = "numeric"      #FATALITIES
importColumns[24] = "numeric"      #INJURIES
importColumns[25] = "numeric"      #PROPDMG
importColumns[26] = "factor"       #PROPDMGEXP
importColumns[27] = "numeric"      #CROPDMG
importColumns[28] = "factor"       #CROPDMGEXP

stormData<-read.csv (dataFileExt, nrows=-1,colClasses = importColumns )

```

- Clean the data
```{r Clean_Data, echo=TRUE}
# We are only interested in data where there was: Property/Crop Damage, 
# Injuries or Fatalities. If all of these data points are 0 for a row we
# can discard it. 

stormData <- stormData[(stormData$PROPDMG > 0 | 
                        stormData$CROPDMG > 0 |
                        stormData$FATALITIES >0 |
                        stormData$INJURIES >0) , ]

#Convert our date's to a POSIX date 
#stormData$BGN_DATE <-as.POSIXct( stormData$BGN_DATE , format = "%m/%d/%Y %H:%M:%S")

# Convert the property damage to actual dollars 
# Estimates should be rounded to three significant digits, followed by an 
# alphabetical character signifying the magnitude of the number, i.e., 1.55B 
# for $1,550,000,000. Alphabetical characters used to signify magnitude
# include “K” for thousands, “M” for millions, and “B” for billions

# Covert the PROPDMGEXP to upper case for consistency
stormData$PROPDMGEXP<- as.factor(toupper(stormData$PROPDMGEXP))
stormData$CROPDMGEXP<- as.factor(toupper(stormData$CROPDMGEXP))

# Filter out rows with bad values
stormData <- stormData[stormData$PROPDMGEXP %in% c("","K","M","B"),]
stormData <- stormData[stormData$CROPDMGEXP %in% c("","K","M","B"),]

#re-level the factor now that rows have been removed
stormData$PROPDMGEXP <-factor(stormData$PROPDMGEXP)
stormData$CROPDMGEXP <-factor(stormData$CROPDMGEXP)
step1Rows <- nrow(stormData)

# There are certain cases where the EXP value was not included but a
# DMG value was associated. Impute these to "K" values, this may lead to 
# under reporting damage estimates
stormData[(stormData$PROPDMG > 0 & stormData$PROPDMGEXP ==""),]$PROPDMGEXP <-"K"
stormData[(stormData$CROPDMG > 0 & stormData$CROPDMGEXP ==""),]$CROPDMGEXP <-"K"

# Add columns for the real $ value based on the *DMG and *DMGEXP columns
expandDollars <- function (dmg,dmgexp) { 
   if (dmg == 0 ) { return(0) }
   else if (dmgexp =="K") {return(dmg*1000)}
   else if (dmgexp =="M") {return(dmg*1000000)}
   else if (dmgexp =="B") {return(dmg*1000000000)}
   }
stormData$PROPDMGDOLLAR<-mapply(expandDollars,stormData$PROPDMG,stormData$PROPDMGEXP)
stormData$CROPDMGDOLLAR<-mapply(expandDollars,stormData$CROPDMG,stormData$CROPDMGEXP)

# Remove data that does not seem to align to a known EVTYPE
# This is bases of off the Events Type list in section 2.1.1 (page 6)
# of the "NATIONAL WEATHER SERVICE INSTRUCTION 10-1605" document

# Remove "Summary","Record","Monthly" rows, these seem to be summarizing data
# and could be leading to double counting 
stormData <- stormData[!grepl("^Summary",stormData$EVTYPE,ignore.case = TRUE),]
stormData <- stormData[!grepl("^Record",stormData$EVTYPE,ignore.case = TRUE),]
stormData <- stormData[!grepl("^Month",stormData$EVTYPE,ignore.case = TRUE),]

# #Rename factors that seem to be miscatagorized - "OTHER" will be the catch all catagory
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="?"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="Other"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="Marine Accident"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MARINE MISHAP"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="APACHE COUNTY"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="DROWNING"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="RAPIDLY RISING WATER"] <- "OTHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HEAVY MIX"] <- "HEAVY RAIN"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="SEVERE TURBULENCE"] <- "HEAVY RAIN"

#Group EVTYPES to align with doc
# Due to overlapping in descriptions the ordering the grouping is done in is fairly sensitive 
 
#TStorm Surge
levels(stormData$EVTYPE)[grepl("SURGE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "STORM SURGE"
 
# #Tide
levels(stormData$EVTYPE)[grepl("TIDE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "ASTRONOMICAL HIGH TIDE"
 
#Avalanche
levels(stormData$EVTYPE)[grepl("AVALA",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "AVALANCHE"

#Blizzard
levels(stormData$EVTYPE)[grepl("BLIZZ",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "BLIZZARD"

#Rip Current
levels(stormData$EVTYPE)[grepl("RIP",levels(stormData$EVTYPE),ignore.case = TRUE)]  <- "RIP CURRENT" 

#High Surf
levels(stormData$EVTYPE)[grepl("SURF",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HIGH SURF"

#Costal Flooding
levels(stormData$EVTYPE)[grepl("COAST",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "COASTALFLUD" #temp

#Debris Flow
levels(stormData$EVTYPE)[grepl("SLIDE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DEBRIS FLOW" 
levels(stormData$EVTYPE)[grepl("SLUMP",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DEBRIS FLOW" 
 
#Extreme
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXTREME HEAT"] <- "EXH1"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXCESSIVE HEAT"] <- "EXH1"
levels(stormData$EVTYPE)[grepl("EXTREME ",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "ECWC1" #temp

#Frost/Freezing
levels(stormData$EVTYPE)[grepl("FROST",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FROST/FREEZE"
levels(stormData$EVTYPE)[grepl("FREEZ",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FROST/FREEZE"

#Fog
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="FOG AND COLD TEMPERATURES"] <-  "DENSE FOG"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="FOG"] <-  "DENSE FOG"
levels(stormData$EVTYPE)[grepl("FREEZING FOG",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FREEZING FOG"
levels(stormData$EVTYPE)[grepl("GLAZE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FREEZING FOG"

#Dust
levels(stormData$EVTYPE)[grepl("SMOKE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DENSE SMOKE"
levels(stormData$EVTYPE)[grepl("DROUGHT",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DROUGHT"
levels(stormData$EVTYPE)[grepl("MICROBURST",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"TWNID" #temp
levels(stormData$EVTYPE)[grepl("DRY",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DROUGHT"
levels(stormData$EVTYPE)[grepl("DUST D",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DUST DEVIL"
levels(stormData$EVTYPE)[grepl("DUST S",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "DUST STORM"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="BLOWING DUST"] <-  "DUST STORM"

#Excessive/Heavy
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXCESSIVE RAINFALL"] <- "HEAVY RAIN"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXCESSIVE SNOW"] <- "HEAVY SNOW"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXCESSIVE WETNESS"] <- "HEAVY RAIN"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HEAVY SHOWER"] <- "HEAVY RAIN"

#Lakes and floods
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LAKE FLOOD"] <- "LAKEFLUD"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LAKESHORE FLOOD"] <- "LAKEFLUD"
levels(stormData$EVTYPE)[grepl("FLOOD",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FLOOD"
levels(stormData$EVTYPE)[grepl("URBAN",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FLOOD"
levels(stormData$EVTYPE)[grepl("STREAM",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FLOOD"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="FLASHFLUD"] <- "FLASH FLOOD"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LAKEFLUD"] <- "LAKESHORE FLOODING"
levels(stormData$EVTYPE)[grepl("COASTALFLUD",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "COASTAL FLOOD"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="GUSTY LAKE WIND"] <- "HIGH WIND"
levels(stormData$EVTYPE)[grepl("LAKE ",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "LESW" #temp
levels(stormData$EVTYPE)[grepl("DAM ",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FLASH FLOOD"

#WaterSpout
levels(stormData$EVTYPE)[grepl("WATERSPOUT",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "WATERSPOUT"

#Funnel Clouds 
levels(stormData$EVTYPE)[grepl("FUNNEL",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "FUNNEL CLOUD"

#Hail (to the king baby)
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MARINE HAIL"] <- "MHIAL"
levels(stormData$EVTYPE)[grepl("HAIL",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "HAIL"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MHIAL"] <- "MARINE HAIL"

#Heat
levels(stormData$EVTYPE)[grepl("HEAT",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "HEAT"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HIGH"] <- "HEAT"
levels(stormData$EVTYPE)[grepl("WARM",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "HEAT"
levels(stormData$EVTYPE)[grepl("HYPERTHERMIA/EXPOSURE",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "HEAT"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="EXH1"] <- "EXCESSIVE HEAT"

#Ice Storm
levels(stormData$EVTYPE)[grepl("ICE STORM",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "ICE STORM"
levels(stormData$EVTYPE)[grepl("ICY",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "ICE STORM"

#Sleet
levels(stormData$EVTYPE)[grepl("SLEET",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "SLEET"

#Cold/Wind Chill
levels(stormData$EVTYPE)[grepl("COLD",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"CWC1" #temp
levels(stormData$EVTYPE)[grepl("CHILL",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"CWC1" #temp
levels(stormData$EVTYPE)[grepl("COOL",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"CWC1" #temp
levels(stormData$EVTYPE)[grepl("LOW TEMP",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"CWC1" #temp
levels(stormData$EVTYPE)[grepl("HYPOTHERMIA",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"CWC1" #temp

#Heavy Snow
levels(stormData$EVTYPE)[grepl("SNOW",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HEAVY SNOW"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LESW"] <-  "LAKE-EFFECT SNOW"

#Rain
levels(stormData$EVTYPE)[grepl("RAIN",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HEAVY RAIN"
levels(stormData$EVTYPE)[grepl("WET",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HEAVY RAIN"
levels(stormData$EVTYPE)[grepl("PRECIP",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HEAVY RAIN"

#Lightening
levels(stormData$EVTYPE)[grepl("LIGHT",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "LIGHTNING"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LIGNTNING"] <-  "LIGHTNING"

#Wind
levels(stormData$EVTYPE)[grepl("MARINE STRONG",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "MSW1" #temp
levels(stormData$EVTYPE)[grepl("MARINE HIGH",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "MSW2" #temp
levels(stormData$EVTYPE)[grepl("strong",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "SWC1"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="NON-TSTM WIND"] <-  "HIGH WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="NON TSTM WIND"] <-  "HIGH WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MARINE TSTM WIND"] <-  "MTWNID" #temp
levels(stormData$EVTYPE)[grepl("THUND",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"TWNID" #temp
levels(stormData$EVTYPE)[grepl("TSTM",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"TWNID"
levels(stormData$EVTYPE)[grepl("WIND",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HIGH WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MTWNID"] <-  "MARINE THUNDERSTORM WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="TWNID"] <-  "THUNDERSTORM WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="CWC1"] <-  "COLD/WIND CHILL"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="ECWC1"] <-  "EXTREME COLD/WIND CHILL"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="SWC1"] <-  "STRONG WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MSW1"] <-  "MARINE STRONG WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="MSW2"] <-  "MARINE HIGH WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="WND"] <-  "WIND"
levels(stormData$EVTYPE)[grepl("GUSTNADO",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"THUNDERSTORM WIND"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="DOWNBURST"] <-  "THUNDERSTORM WIND"

# High Surf
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HIGH SWELLS"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HIGH WATER"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HIGH SEAS"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HIGH WAVES"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[grepl("beach",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "HIGH SURF" 
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HEAVY SEAS"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="HEAVY SWELLS"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="ROGUE WAVE"] <-  "HIGH SURF"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="ROUGH SEAS"] <-  "HIGH SURF"

#Hurricane 
levels(stormData$EVTYPE)[grepl("HUR",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HURRICANE/TYPHOON"
levels(stormData$EVTYPE)[grepl("TYP",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"HURRICANE/TYPHOON"

#Ice Storm
levels(stormData$EVTYPE)[grepl("ICE",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"ICE STORM"

#Tornado
levels(stormData$EVTYPE)[grepl("TORNADO",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"TORNADO"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="TORNDAO"] <-  "TORNADO"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="LANDSPOUT"] <-  "TORNADO"

#Topical Storm
levels(stormData$EVTYPE)[grepl("TROPICAL ST",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"TROPICAL STORM"

#Wildfire
levels(stormData$EVTYPE)[grepl("FIRE",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"WILDFIRE"

#Winter Weather and Winter Storm
levels(stormData$EVTYPE)[grepl("WINTER ST",levels(stormData$EVTYPE),ignore.case = TRUE)] <-"XWNT" #temp
levels(stormData$EVTYPE)[grepl("WINT",levels(stormData$EVTYPE),ignore.case = TRUE)] <- "WINTER WEATHER"
levels(stormData$EVTYPE)[levels(stormData$EVTYPE)=="XWNT"] <-  "WINTER STORM"

#stormData$EVTYPE <-factor (stormData$EVTYPE)
```

## Results
- The analysis document must have at least one figure containing a plot.
- Your analyis must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total

### Question 1: Across the United States, which types of events are most harmful with respect to population health
```{r Question_One, echo=TRUE, fig.height=10, fig.width=12, message=FALSE}
require(ggplot2)
require(reshape2)
require(scales)

#Question 1: Across the United States, which types of events are most harmful with respect to population health
totalsByEventType<- aggregate( cbind(FATALITIES,INJURIES) ~ EVTYPE, data=stormData, FUN=sum)
discreteEventInjuries<-melt(totalsByEventType, id.vars = c("EVTYPE"),measure.vars = c("FATALITIES","INJURIES") )
p = ggplot(data=discreteEventInjuries, aes(x=reorder(EVTYPE,value), y=value,fill=variable)) +
   scale_fill_discrete(name="Type",
                       breaks=c("FATALITIES","INJURIES"),
                       labels=c("Fatalities","Injuries")) +
   geom_bar(stat="identity") +
   theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = .5)) +
   scale_y_sqrt(labels=comma ) +
   xlab("Event Type") +
   ylab("Fatalities + Injuries") +
   ggtitle("Fatalities and Injuries by Storm Event Type")

print (p)
```

### Question 2: Across the United States, which types of events have the greatest economic consequences
```{r Question_two, echo=TRUE,fig.height=10, fig.width=12}
#Question 2: Across the United States, which types of events have the greatest economic consequences
totalsByEventType<- aggregate( cbind(PROPDMGDOLLAR,CROPDMGDOLLAR) ~ EVTYPE, data=stormData, FUN=sum)
discreteEventDamage<-melt(totalsByEventType, id.vars = c("EVTYPE"),measure.vars = c("PROPDMGDOLLAR","CROPDMGDOLLAR") )
p = ggplot(data=discreteEventDamage, aes(x=reorder(EVTYPE,value), y=value,fill=variable)) +
   scale_fill_discrete(name="Type",
                       breaks=c("PROPDMGDOLLAR","CROPDMGDOLLAR"),
                       labels=c("Property Damage","Crop Damage")) +
   geom_bar(stat="identity") +
   theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = .5)) +
   scale_y_sqrt(breaks=c(1000000000,5000000000,10000000000,50000000000,100000000000,200000000000),labels=dollar ) +
   xlab("Event Type") +
   ylab("Damage in Dollars") +
   ggtitle("Damage by Event Type")


print(p)

```